
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

```{r}
library(tidyverse)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
head(df)
```
Hint: 

-  Use `filter` to select the most current date

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
#The countries with the highest number of cumulative deaths were the U.S., Brazil, and India.

df2 <- df %>% 
  filter(Date_reported == max(Date_reported)) %>%
  arrange(-Cumulative_deaths)
head(df2)
  
```

2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths

```{r}
library(ggplot2)
library(knitr)
library(dplyr)
library(gganimate)
library(gifski)

df2 %>% 
  filter(Country_code == "US"|
           Country_code == "BR"|
           Country_code == "IN") %>% 
  ggplot(aes(x=Country_code,y=Cumulative_deaths))+
  transition_reveal(along = Cumulative_deaths)
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df <- df %>%
  mutate(death_per_cases = Cumulative_deaths/Cumulative_cases)
df
```


4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 

```{r}

```

5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv

```{r}

```

Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot.

```{r}

```

