
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Data Wrangging with `dplyr`"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 4. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so. 


-------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://covidtracking.com/data/download/all-states-history.csv).  Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv.

```{r}
library(tidyr)
library(tidyverse)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```


-------

#### 3.  Fix the date and ceate some new variables

`lubridate` is a package of the `tidyverse` packages.  We will make uses of `lubridate` in this question.

- Use the below codes to create `month`, `weekday` and `monthday` variables

```{r}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
head(df)
```

-------

#### 4. Create new variables with `case_when`. 

The function `case_when` is a good option to create a new variable from existing variable.  For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable.  `deathIncrease` is the number of daily new death by Covid19.  The new variable `daily_death` takes three values:  low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14).  Please notice that this can also be done in a different way as shown in Assignment 3. 

```{r}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
)
head(df)
```

- Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month > 20).

- Create variable `weekend` that takes two values:  1 if it's Saturday or Sunday or 0 otherwise.

```{r}
df$month2 <- case_when(
  df$monthday <11 ~ 'early_month',
  df$monthday <=20 ~ 'mid_month',
  TRUE ~ 'end_month'
)
head(df)
```
```{r}
df$weekend <- case_when(
  df$weekday <2 ~ 1,
  df$weekday <=6 ~ 0,
  TRUE ~ 1
)
head(df)
  
```


-------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.

```{r}
#Deselecting totalTestsViral with the select function.
#The deselection was assigned to the dataframe.

df <- df %>% select(-totalTestsViral)
df

```


-------

#### 6. Pipe Operator ( %>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable.  Pipe works very well with all the `tidyverse` packages. Refer to [these slides](../slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
sqrt(x)

sum(sqrt(x))

log(sum(sqrt(x)))

# log base 2 of 16
log(16, 2)

```
```{r}
x %>% sqrt
```
```{r}
x %>% sqrt %>% sum
```
```{r}
x %>% sqrt %>% sum %>% log

```
```{r}

16 %>% log2

```

-------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

```{r}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) by `month`

- Calculate the average number of cases (`positiveIncrease`) by `month2`

- Calculate the median number of cases (`positiveIncrease`) by `weekend`
```{r}
#Median number of cases by 'month'

df %>% 
  group_by(month) %>% 
  summarise(median(positiveIncrease))

```
```{r}
#Average number of cases by 'month2'

df %>% 
  group_by(month2) %>% 
  summarise(mean(positiveIncrease))
```
```{r}
#Median number of cases by 'weekend'

df %>% 
  group_by(weekend) %>% 
  summarise(median(positiveIncrease))
```


-------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020. 
```{r}
df %>% 
  filter(month==10|month==11,year(date)==2020) %>% 
  group_by(weekend) %>% 
  summarise(positve_increase = median(positiveIncrease))
```


- Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020
```{r}
df %>% 
  filter(year(date) == 2020 & ((month == 9 & monthday >21) | month == 10 |
                                 month == 11 | (month == 12 & monthday <= 20))) %>% 
  group_by(month2) %>% 
  summarise(mean_death = mean(death, na.rm = TRUE))
```
- Compare the average number of hospitalizations between weekdays and weekends in Summer 2020
```{r}
df %>% 
  filter(year(date) == 2020 & ((month == 6 & monthday > 19) | month == 7 | 
                                 month == 8 |(month == 9 & monthday <21))) %>% 
  group_by(weekend) %>% 
  summarise(mean_hospitalized = mean(hospitalized, na.rm = TRUE))
```

- Redo Questions 14 and 15 in [Assignment 3](assignment3.html) using the combos.  Notice:  you also need to use the data used in Assignment 3.

14. Find the median cases by weekdays in Rhode Island in 2021
```{r}
df %>% 
  filter(state == 'RI' & year(date) == 2021) %>% 
  group_by(weekday) %>% 
  summarise(mean_cases = mean(positiveIncrease, na.rm = TRUE))
```
15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
df %>% 
  filter(state == 'RI' & year(date)==2021 & month == 6 | month == 7 | month == 8 | month == 9) %>% 
  group_by(month) %>% 
  summarise(median_cases = median(positiveIncrease, na.rm = TRUE))
  
```


-------

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend. 
```{r}
df %>% filter(weekend == 1) %>% 
  group_by(month) %>%
  summarise(deaths_by_weekend = sum(death, na.rm = TRUE)) %>% 
  arrange(-deaths_by_weekend)


```


-------

#### 10. Use your own dataset and implement the follows functions or combos. You can use the [Adult Census Income](../data/adult_census.csv) or [Titanic data](../data/titanic.csv). 

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange

```{r}
#Importing a data set on universities. source (https://www.kaggle.com/datasets/alitaqi000/world-university-rankings-2023)

universityDf <- read_csv('World University Rankings 2023.csv')
universityDf
```
```{r}
#Removing the column 'No of student per staff' from the dataframe.

universityDf <- universityDf %>% select(-'No of student per staff')
universityDf
```
```{r}
universityDf %>% filter(Location == 'United States')
```
```{r}
#Converting the class of variables in the dataframe from character to numeric with mutate.

universityDf <- universityDf %>% mutate(`OverAll Score`=as.numeric(`OverAll Score`),
                        `Teaching Score`=as.numeric(`Teaching Score`),
                        `Research Score`=as.numeric(`Research Score`),
                        `Citations Score`=as.numeric(`Citations Score`),
                        `Industry Income Score`=as.numeric(`Industry Income Score`),
                        `International Outlook Score`=as.numeric(`International Outlook Score`), na.rm =TRUE)
```
```{r}
universityDf
```
```{r}
universityDf %>% summarise(min_number_of_students = min(.[[4]], na.rm = TRUE),
                           avg_number_of_students = mean(.[[4]], na.rm = TRUE),
                           max_number_of_students = max(.[[4]], na.rm = TRUE))
```
```{r}
#Using subset to find the schools with the minimum and maximum number of students.

universityDf %>% subset(.[[4]] == 115 |
                          .[[4]] == 460632)
```
```{r}
#Using the arrange function to find the universities with the most students.

universityDf %>% arrange(-.[[4]])
```
```{r}
#Using count to find the number of universities in each country.

universityDf %>% count(Location)
```
```{r}
#Using the count + arrange combo to find the locations with the highest number of universities.

universityDf %>% count(Location) %>% 
  arrange(-n)
```
```{r}
#Filter + count + arrange combo: Comparing the number of top small universities (less than 5,000 students)
#between the U.S. and U.K.

universityDf %>% filter(.[[4]]< 5000 & (Location == 'United States'|Location =='United Kingdom')) %>% 
  count(Location) %>% 
  arrange(-n)

```
```{r}
#group_by + summarise

universityDf %>%
  group_by(Location) %>% 
  summarise(mean_ovr_score = mean(`OverAll Score`,na.rm=TRUE))

```
```{r}
#Using the filter + group_by + summarise combo to compare the average overall score between the U.S. and the U.K.

universityDf %>% 
  filter(Location == 'United States'|Location == 'United Kingdom') %>% 
  group_by(Location) %>% 
  summarise(mean_ovr_score = mean(`OverAll Score`,na.rm=TRUE))
```
```{r}
#Using the filter + group_by + summarise + arrange combo to find the countries with the highest mean over all scorem among the world's best universities.

universityDf %>% 
  filter(.[[4]] >= 5000 & .[[4]] <= 15000) %>% 
  group_by(Location) %>% 
  summarise(mean_ovr_score = mean(`OverAll Score`,na.rm=TRUE)) %>% 
  arrange(-mean_ovr_score, na.rm = TRUE)
```

