
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

-------

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------

1. Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)

```{r}
#Importing libraries and data. Source: https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)

df <- read_csv('Billionaires Statistics Dataset.csv')
df
```

```{r}
#When comparing the richest man with the richest woman, they are 74 and 69 years old. The richest people in the world are mostly seniors between 60 to 80. Furthermore, this graph indicates that there are far more male billionaires. The top 10 richest people in the world are all male, and there are more points in the graph on the right. Note (the finalWorth is in millions)

df %>% ggplot()+ 
    geom_point(mapping=aes(x=age, y= finalWorth))+
  facet_wrap(~gender)+
  labs(title = "Billionaires by Gender",
       caption = "The net worth of billionaires are compared by age across gender.",
      tag = "Figure 1",x = "Age",y = "Final Worth")
```

2. Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)

```{r}
#Converting the class of selfMade from logical to character.

df$selfMade <- df$selfMade %>% ifelse('True','False')
df %>% head %>% select(selfMade)

```


```{r}
#Using ggplot's facet_wrap function, the majority of male billionaires appear to be self made while the majority of female billionaires are not self made. The world's richest man and the world's richest woman are not self made. For some more detail on the variable `selfMade`, it is defined as someone who achieved their wealth without a large inheritance of wealth or status.

df %>% ggplot()+ 
  geom_point(mapping=aes(x=age, y=finalWorth ))+
  facet_wrap(~gender+selfMade)+
  labs(title = "Self Made Billionaires by Gender",
       caption = "The final net worth is compared across gender and between
       whether they where self made and age.",
      tag = "Figure 2",x = "Age",y = "Final Worth")
```

3. Make a plot using `position ='dodge'`.

```{r}
#The argument `position = 'dodge'' is best used for comparing graphs of data side by side rather than overlapping. In comparison to fill, dodge makes it easier to compare quantities of variable categories.The most applicable plot to make for either fill or dodge is a bar plot. When comparing the industries of male and female billionaires in the U.S., the industries with the most male billionaires are finance & investments, and technology. The industries with the most female billionaires are food & beverage, fashion & retail, and technology. 

df %>% filter(country == 'United States') %>% ggplot()+ 
  geom_bar(mapping=aes(x=gender, fill=industries), 
           position = 'dodge')+
  labs(title = "Billionaires by Industry",
       caption = "The number of billionaires are compared 
       between industries and gender.",
      tag = "Figure 3",x = "Gender",y = "Number of Billionaires")

```


4. Make a plot using `position ='fill'`. 

```{r}
#The barplot for self made billionaires by gender gives a much better illustration than the second graph of the proportion of self made billionaires between males and females.

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=gender, fill=selfMade), 
           position = 'fill')+
  labs(title = "Self Made Billionaires by Gender",
       caption = "The proportion of self made billionaires
       are compared across gender.",
      tag = "Figure 4",x = "Gender",y = "Proportion")

```


5. Make a plot using `geom_col`.

```{r}
#Interestingly, there is a very high number of billionaires born in the month of January in comparison to the other months. This behavior is strange since the number of billionaires born in January doubles even the second highest month. 

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=birthMonth))+
  labs(title = "Billionaires by Birth Month",
       caption = "A bar plot is used to represent the number
       of billionaires by birth month.",
      tag = "Figure 5",x = "Birth Month",y = "Number of Billionaires")

```


6. Change the theme of one of your plots. 

```{r}
#Library ggthemes was imported to access more themes. `theme_economist()` was used at the end to change the theme. The arguments 'fill' and color were used to change the color of the bars and the outline in the graph.

library(ggthemes)

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=birthMonth), fill = 'yellow',
           color = 'black')+
  labs(title = "Billionaires by Birth Month",
       caption = "A bar plot with extra detail 
       is used to represent the
       number of billionaires by birth month.",
      tag = "Figure 6",x = "Birth Month",y = "Number of Billionaires",
    )+
  theme_economist()
```


7. Make a plot using combo filter + ggplot

```{r}
#Using the filter + ggplot combo the plot the final net worth of billionaires by age in the technology, finance & investments, and manufacturing industries. There is yet to be a person from one of these industries to reach a net worth of $100 billion USD.

df %>% filter(industries == 'Technology'|
                industries == 'Finance & Investments '|
                industries == 'Manufacturing') %>% 
  ggplot()+
  geom_point((aes(x=age, y=finalWorth)), color = 'darkgreen')+
  labs(title = "Billionaires in 3 Industries",
       caption = "Billionaires from the Technology, 
       Finance & Investments, and Manufacturing sectors are compared",
      tag = "Figure 7",x = "Age",y = "Final Worth")+
  theme_igray()
```


8. Make a plot using combo group_by + summarise + geom_col

```{r}
#Using the group_by + summarise + geom_col combo to plot the average net worth by industry among billionaires based on whether their wealth was selfmade. For self made billionaires, the average net worth was highest for those in the automotive industry. On the other hand, the average net worth was highest for billionaires who were not self made and are from the finance & investments industry.

df %>% group_by(industries, selfMade) %>% 
  summarise(mean_worth = mean(finalWorth, na.rm = TRUE)) %>% 
  ggplot()+
  geom_col(aes(x=selfMade, y=mean_worth, fill=industries), position = 'dodge')+
  labs(title = "Average Net Worth by Industry",
       caption = "The average networth of a billionaire is compared by industry",
      tag = "Figure 8",x = "selfMade",y = "Average Final Worth")
  
```


9. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
[Link to the plot](abc.png)
```

```{r}
#Ranking the countries by the highest number of billionaires.

df$country %>% table %>%
  sort(decreasing = TRUE) %>% 
  names
```

```{r}
#Making a plot to save.

plot <- df %>% filter(country == "United States"|
                country == "China"| country == "India"|
                country == "Germany"|
                country == "United Kingdom"|
                country =="Russia"|country == "Switzerland"|
                country == "Hong Kong"| country == "Italy"|
                country == "Singapore") %>% 
  group_by(country,gender) %>%
  summarise(mean_finalWorth = mean(finalWorth, na.rm = TRUE)) %>% 
  ggplot()+
  geom_col(aes(x = country, y = mean_finalWorth, fill = gender), position = 'dodge')+
  labs(title = "Billionaires in 10 Countries",
       caption = "The average net worth of a billionaire is compared by gender in ten countries.",
      tag = "Figure 9",x = "Country",y = " Average Final Worth")+
    theme_fivethirtyeight()
plot
```
```{r}
#Saving the plot.

ggsave(filename = 'billionaires.png', plot = plot)
```

10. In this question, we will use  the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv)

- [Alternative Link](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv')
```



`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))
```

Do the follows. 

  - Use `geom_point` to compare the vaccines used by date in Argentina in 2022. 
  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 
  
```{r}
#Filtering by location and year then plotting the total vaccinations by date. The two following graphs compare 6 different types of vaccines used in Argentina in 2022. The most popular vaccine was 'Sinopharm/Beijing while Moderna and Pfizer/BioNTech increased the most in total vaccinations over that year.

df %>% 
  filter(location=='Argentina', year(date) == '2022') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
  labs(title = "Types of Vaccines Used in Argentina",
       caption = "The number of vaccinations in Argentina in 2022 
       were compared across 6 vaccines.",
      tag = "Figure 10",x = "Date",y = "Total Vaccinations")+
    theme_fivethirtyeight()
```
```{r}
#Filtering by location and year then plotting the total vaccinations by date. This graph is similar to the previous one, but there's a line connecting the observations.

df %>% 
  filter(location=='Argentina', year(date) == '2022') %>% 
  ggplot()+
  geom_line(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
  labs(title = "Types of Vaccines Used in Argentina",
       caption = "The number of vaccinations in Argentina in 2022 
       were compared across 6 vaccines.",
      tag = "Figure 11",x = "Date",y = "Total Vaccinations")+
    theme_fivethirtyeight()
```

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Notice that the above codes could be piped in one 

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```


Do the follows. 

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021

```{r}
#Viewing the locations within the dataset.

table(df$location)
```


```{r}
#Cyprus was not included since it is located on the Mediterranean between Europe and Asia, but it is listed as a European country. In comparison, to the other countries/regions in Asia, Japan has the highest number of vaccinations by a massive margin.

df %>% 
  filter(location %in% c('Hong Kong', 'Japan', 'Nepal', 'South Korea')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
  labs(title = "Total Vaccinations by Country/Region in Asia",
       caption = "The number of vaccinations in are compared across four
       countries / regions in Asia.",
      tag = "Figure 11",x = "date",y = "total_vaccinations")
  
```
```{r}
#Filtering by year and location, then using group_by and summarise to compare the total vaccines by South American nation. In 2021, Argentina and Chile had increased their total vaccines the most out of any country. Uruguay had few vaccinations compared to the other countries, but it has the smallest population. 

df %>% 
  filter(year(date) == '2021', location %in% c('Argentina', 'Chile', 'Ecuador', 'Peru' ,'Uruguay')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
  labs(title = "Total Vaccinations by South American Nation",
       caption = "The number of vaccinations are compared
       across five South American nations.",
      tag = "Figure 12",x = "Date",y = "Total Vaccinations")
```
```{r}
#(Optional) If you wanted to go into more detail and compare the different types of vaccines across countries, ggplot's `facet_wrap()` function can be applied at the end. This example was done for South American Countries.

df %>% 
  filter(location %in% c('Argentina', 'Chile', 'Ecuador', 'Peru' ,'Uruguay')) %>% 
  group_by(date, location, vaccine) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=vaccine))+facet_wrap(~`location`)+
  labs(title = "Vaccines by South American Nation",
       caption = "The number of vaccinations in Argentina in 2022 
       were compared across 6 vaccines and 6 types of vaccines.",
      tag = "Figure 13",x = "Date",y = "Total Vaccinations")+
  theme_economist()

```