
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

```{r}
df = data.frame(Rank = c('0','1','2','3'),
                Age = c(28,34,29,42),
                Name = c('Tom','Jack','Steve',
                         'Ricky'))
                         
df
```


2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

```{r}
df2 <- read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
head(df2)
```


2. Show the names of the variables in the data

```{r}
names(df2)
```


3. How many columns and rows the data have? 

```{r}
#There are 5 columns and 61,942 rows.
summary(df2$state)
str(df2)
```


4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
sum(is.na(df2))

colSums(is.na(df2))

# There are no missing values; therefore, all
# variables have no missing values.

```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 

```{r}
class(df2$date)

df2$date = as.Date(df2$date)

class(df2$date)

# The class of the date column was converted from
# character to date.
```


6. Capitalize the names of all the variables

```{r}

names(df2)[1] <- 'Date'
names(df2)[2] <- 'State'
names(df2)[3] <- 'Fips'
names(df2)[4] <- 'Cases'
names(df2)[5] <- 'Deaths'
head(df2)


```


7. Find the average number of cases per day. Find the maximum cases a day.

```{r}

mean(df2$Cases, na.rm = TRUE)
max(df2$Cases, na.rm = TRUE)

```

8. How many states are there in the data?  

```{r}
#The data includes all 50 states and 6 U.S. Territories.

names(table(df2$State))
```


9. Create a new variable `weekdays` to store the weekday for each rows.

```{r}

df2$weekdays <- weekdays(df2$Date)
head(df2)

```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`. 
```{r}
df2$death2 <- ifelse(df2$Deaths==0,
                     'no_death','has_death')

head(df2)
```

```{r}
#Using R to calculate frequency and relative frequency (method 1)
no_death = sum(df2$death2 == 'no_death')

has_death = sum(df2$death2 == 'has_death')

table(df2$death2)

prop.table(table(df2$death2))

```

```{r}
#Since this was answered in the cell above. Relative frequency was
#calculated using different method in this cell.

print("Relative frequency of no_death")

no_death/(no_death+has_death)

print("Relative frequency of has_death")

has_death/(no_death+has_death)

```


11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df2$Deaths)
```
12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)

df2$death3 <- case_when(df2$Deaths<1171 ~ 'low_death',
                             df2$Deaths<14912 ~ 'mid_death',
                             TRUE~'high_death')
table(df2$death3)

```
```{r}
#Viewing the first 6 rows of the dataframe.
head(df2)
```


13. Find the average cases in Rhode Island in 2021

```{r}
#Subsetting the dataframe based on cases in 2021.

df3 <- df2[(df2$Date<='2021-12-31')&
           (df2$State == 'Rhode Island'),]

df3 <- df3[df3$Date>='2021-01-01',]
head(df3)

```
```{r}
#This line calculates the average number of cases in 2021. 
#The average number of cases in Rhode Island in 2021 is 154,439.

summary(df3$Cases)
```


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
#Subsetting the dataframe by weekday and taking the median of cases for each subset.


by(df3$Cases, df3$weekdays, median)


```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 

```{r}
#Creating subset dataframes and comparing the median cases. The median number of cases has increased from June through September of 2021.

dfJune <- df3[(df3$Date <= '2021-06-30')&
                 (df3$Date>='2021-06-01'),]

dfJuly <- df3[(df3$Date <= '2021-07-31')&
                 (df3$Date>='2021-07-01'),]

dfAugust <- df3[(df3$Date <= '2021-08-31')&
                 (df3$Date>='2021-08-01'),]

dfSeptember <- df3[(df3$Date <= '2021-09-30')&
                 (df3$Date>='2021-09-01'),]

print("June")
median(dfJune$Cases)
print("July")
median(dfJuly$Cases)
print("August")
median(dfAugust$Cases)
print("September")
median(dfSeptember$Cases)
```
16. Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

```{r}
#Importing a dataset from Kaggle: Source (https://www.kaggle.com/datasets/justinas/nba-players-data?resource=download)

#The head function was used to display the first 6 rows of the data rather than the entire dataset.

nbaDf <- read.csv('nba.csv')

head(nbaDf)
```

```{r}
str(nbaDf)
```

```{r}
names(nbaDf)
```

```{r}
averageHeight = mean(nbaDf$player_height)
minHeight = min(nbaDf$player_height)
maxHeight = max(nbaDf$player_height)

```

```{r}
averageHeight
```
```{r}
minHeight
```
```{r}
maxHeight
```
```{r}
#Converting height in centimeters to inche. (This was not used on the dataframe)
minHeight = minHeight/cm(1)
maxHeight = maxHeight/cm(1)
averageHeight = averageHeight/cm(1)
```
```{r}
averageHeight
```
```{r}
minHeight
```
```{r}
maxHeight
```
```{r}
table(nbaDf$team_abbreviation)
```
```{r}
sum(is.na(nbaDf))
```
```{r}
colSums(is.na(nbaDf))
```
```{r}
class(nbaDf$player_weight)
      
```
```{r}
cor(nbaDf$player_height,nbaDf$player_weight)
```

```{r}
#Viewing the average height of NBA players by country since 2000.
nbaDf1 <- nbaDf[nbaDf$season>='2000-01',]
by(nbaDf1$player_height, nbaDf1$country, mean)
```

```{r}

head(ifelse(nbaDf$player_height > cm(84), "Over 7 Feet.",
       "Under 7 Feet"))
```
```{r}
#Using the case_when function to find out how many nba players fall into each category of rebounding.

rebounding <- case_when(nbaDf$reb<2.5 ~ 'poor_rebounder',
                             nbaDf$reb<8 ~ 'average_rebounder',
                             TRUE~'great_rebounder')
table(rebounding)
```

17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories

```{r}
#Creating a categorical variable from the continuous variable height.

nbaDf$Seven_Feet <- ifelse(nbaDf$player_height > cm(84), "Over 7 Feet.",
                    "Under 7 Feet.")
head(nbaDf)
```
```{r}
#Creating a new categorical variable from country.

playerInfo <- ifelse(nbaDf$country == "USA",'Local Player','Foreign Player')
nbaDf$Player_Info <- playerInfo
head(nbaDf)
```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
