
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7 - Extra Credits: Practice dplyr and ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas



```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


-------

1. Import the data at https://bryantstats.github.io/math421/data/ds_salaries.csv

    The data description can be found at https://www.kaggle.com/datasets/saurabhshahane/data-science-jobs-salaries
    
```{r}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)

df <- read_csv('https://bryantstats.github.io/math421/data/ds_salaries.csv')
df
```


2. Is there any column that contains NAs?
```{r}
#There are no columns that contain NAs.
colSums(is.na(df))
```
3. What is the most frequent job title in the data?

```{r}

df$job_title %>% table %>% 
  which.max %>% names

df$job_title %>% table %>% 
  max

```
4. Find the average salary data of the three most frequent job titles in the data. 
```{r}
df$job_title %>% table %>% 
  sort(decreasing = TRUE)
```
```{r}
df %>% 
  filter(job_title == c('Data Scientist', 'Data Engineer', 'Data Analyst')) %>%
  group_by(job_title) %>% 
  summarise(mean_job_salary = mean(salary))
```


```{r}
dsDf <- df %>% filter(job_title == 'Data Scientist')
mean(dsDf$salary)
```
```{r}
df %>% filter(job_title == 'Data Engineer') %>% 
  summarise(mean_salary_eng = mean(salary))
```
```{r}
df %>% filter(job_title == 'Data Analyst') %>% 
  summarise(mean_salary_da = mean(salary))
```
5. What is the job title that have the highest average salary?

```{r}
#Head of Machine Learning.

df %>% 
  group_by(job_title) %>% 
  summarise(avg_job_salary = mean(salary)) %>% 
  arrange(-avg_job_salary)
```


6. Which country gives the highest salary on average? 

```{r}
#CL
df %>% group_by(company_location) %>% 
  summarise(avg_country_salary = mean(salary)) %>% 
  arrange(-avg_country_salary)
```


7. Plot five meaningful plots. All plots should have title, caption, and appropriate labels.

```{r}
#Tracking the growth in average salary from 2020 to 2022.
df %>%
  group_by(work_year) %>% 
  summarise(mean_job_salary = mean(salary_in_usd)) %>% 
  ggplot(aes(x=work_year, y = mean_job_salary))+
  geom_col()+
  labs(title = "Average Salary", 
       caption = "The average salary 
       is plotted from 2020 to 2022.", 
       tag = "Figure 1",x = "Work Year", y = " Avg. Salary")
  
  
```
```{r}
#Plotting the average salary by experience level.

df %>%
  group_by(experience_level) %>% 
  summarise(mean_salary = mean(salary_in_usd)) %>% 
  ggplot(aes(x=experience_level,y=mean_salary))+
  geom_col()+
  labs(title = "Average Salary by Experience Level", 
       caption = "The average salary 
       is compared across all levels of experience", 
       tag = "Figure 2",x = "Experience Level", y = " Avg. Salary")
  
  
```
```{r}
#Plotting the increase in the average proportion of workers that work remote.

df %>% 
  group_by(work_year) %>% 
  summarise(avg_remote_ratio = mean(remote_ratio)) %>% 
  ggplot(aes(x=work_year,y=avg_remote_ratio))+
  geom_line()+
  labs(title = "Average Remote Ratio by Year", 
       caption = "The average remote ratio is compared by year", 
       tag = "Figure 3",x = "Year", y = " Avg. Remote Ratio")
  
  
```
```{r}
#Plotting the average salary in USD by country.

df %>%
  group_by(company_location) %>% 
  summarise(avg_salary = mean(salary_in_usd)) %>% 
  ggplot(aes(x=company_location,y=avg_salary, color = company_location))+ 
  geom_col()+
  labs(title = "Average Salary", 
       caption = "The average salary 
       is compared across company locations", 
       tag = "Figure 4",x = "Country", y = " Avg. Salary")
  
```
```{r}
df %>% 
  group_by(company_location,company_size) %>%
    summarise(mean_salary = mean(salary)) %>% 
    ggplot()+
    geom_col(aes(x = company_location, y = mean_salary, fill = company_size), position = 'dodge')+
    labs(title = "Average Salary",
         caption = "The average salary is compared by company size
         among the company locations.",
        tag = "Figure 5",x = "Country",y = " Avg. Salary")
  
```

