
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
install.packages('tidyr')
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.  
```{r}
library(tidyverse)
df <- read_csv('https://bryantstats.github.io/math421/data/adult_census_missing.csv')
head(df)
```
```{r}
colSums(is.na(df))
```
```{r}
sum(is.na(df))
```


2. Using function `aggr` the VIM package to plot the number of NAs for each column. 

```{r}
install.packages('VIM')
```

```{r}
library('VIM')
aggr(df)
```

3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 

```{r}
table(df$age)
```
```{r}
#There are suspicious categories in the variable 'workclass'. There categories include
#'?', 'Not known', 'Unknown',
table(df$workclass)
```


4. Replace all the forms of missing values found to NA

```{r}
df1 <- df %>% 
  replace('?','Unknown', NA)
```


5. Replot the number of NAs for each column. 

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 





