
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.  
```{r}
#Importing the tidyverse library and dataset with read_csv.

library(tidyverse)
df <- read_csv('https://bryantstats.github.io/math421/data/adult_census_missing.csv')
head(df)
```
```{r}
#Checking for missing values by column.
colSums(is.na(df))
```
```{r}
#counting the number of NAs.
sum(is.na(df))
```


2. Using function `aggr` the VIM package to plot the number of NAs for each column. 

```{r}
#Importing VIM and plotting missing values.

library('VIM')
aggr(df)
```

3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
#Listing the variable names.
names(df)
```


```{r}
#There are no suspicious categories within the variable 'age'.
table(df$age)
```
```{r}
#There are suspicious categories in the variable 'workclass'. There categories include
#'?', 'Not known', and 'Unknown'.
table(df$workclass)
```
```{r}
#There are no suspicious categories within the variable 'fnlwgt'.

head(table(df$fnlwgt))
max(df$fnlwgt)
min(df$fnlwgt)
```
```{r}
#There are suspicious categories in the variable education. These categories are
#'Not Known' and 'Unknown'
table(df$education)
```
```{r}
#There are no suspicious categories within the variable 'education.num'.
table(df$education.num)
```
```{r}
#There is one suspicious category within the variable 'marital.status'. This category
#is titled 'Unknown'.
table(df$marital.status)
```
```{r}
#There are 3 suspicious categories within the variable 'occupation.'
#These categories include '?','Not Known', and 'Unknown'
table(df$occupation)
```
```{r}
#There are two suspicious category within the variable 'relationship'.
#The category is titled 'Unknown' and 'Not Known'
table(df$relationship)
```
```{r}
#There are no suspicious categories within the variable race.
table(df$race)
```

```{r}
#There is one suspicious category within the variable 'sex'.
#The category is titled 'Unknown'.

table(df$sex)
```
```{r}
#There are no suspicious categories within the variable 'capital.gain'.
table(df$capital.gain)
```
```{r}
#There are no suspicious categories within the variable 'capital.loss'.
table(df$capital.loss)
```
```{r}
#There are no suspicious categories within the variable 'hours.per.week'.
table(df$hours.per.week)
```
```{r}
#There are 3 suspicious categories within the variable 'native.country.'
#These categories include '?','Not Known', and 'Unknown'.
table(df$native.country)
```
```{r}
#There are no suspicious categories within the variable 'income'.
table(df$income)
```

4. Replace all the forms of missing values found to NA

```{r}
#Replacing all forms of missing values to NA within the dataframe. Since the same suspicious
#categories have appeared in multiple variables, it is most efficient to us mutate_all.

df1 <- df %>% mutate_all(~replace(., . == 'Not Known', NA)) %>% 
  mutate_all(~replace(., . == '?', NA)) %>% 
  mutate_all(~replace(., . == 'Unknown', NA))

df1
```


5. Replot the number of NAs for each column. 

```{r}
#As expected, the number of Nas by column has increased sharply after converting categories
#to NA.
colSums(is.na(df1))
```
```{r}
#Plotting the number of missing values by column.
aggr(df1)
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?

```{r}
#There are 29,865 rows left after removing the missing values by row.
drop_na(df1)
```


7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
#Checking for missing values by column.

colSums(is.na(df1))

```


```{r}
#Replacing NAs with the next value in the columns with missing values. The changes are saved to a new dataframe.
df2 <- df1 %>% fill(age, workclass, education, 
                    marital.status, occupation,
                    relationship, sex, capital.gain, native.country,
                    .direction = c("updown"))
colSums(is.na(df2))
```
```{r}
#Checking the dataframe.
head(df2)
```

8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
#In the dataframe, there are 9 variables with missing values. Of these variables, 2 are numeric. Therefore the columns that will be operated on are 'age' and 'capital.gain'.
df1 %>% select_if(is.numeric)
```
```{r}
#creating variables for the median of the two columns, and replacing the missing values with the median of each corresponding column.
median_age = median(df1$age, na.rm = TRUE)
median_capital.gain = median(df1$capital.gain, na.rm = TRUE)
df1$age <- replace_na(df1$age, median_age)
df1$capital.gain <- replace_na(df1$capital.gain, median_capital.gain)
colSums(is.na(df1))

```
```{r}
#Finding the majority category for each of the 7 categorical variables with missing values.
df1 %>% select_if(is.character)
```
```{r}
#Assigning variables to the most frequent class that appears within each categorical variable. The changes are then saved to the dataframe.

majority_workclass = names(which.max(table(df1$workclass)))
majority_education = names(which.max(table(df1$education)))
majority_marital.status = names(which.max(table(df1$marital.status)))
majority_occupation = names(which.max(table(df1$occupation)))
majority_relationship = names(which.max(table(df1$relationship)))
majority_sex = names(which.max(table(df1$sex)))
majority_native.country = names(which.max(table(df1$native.country)))

df1$workclass <- replace_na(df1$workclass, majority_workclass)
df1$education <- replace_na(df1$education, majority_education)
df1$marital.status <- replace_na(df1$marital.status, majority_marital.status)
df1$occupation <- replace_na(df1$occupation, majority_occupation)
df1$relationship <- replace_na(df1$relationship, majority_relationship)
df1$sex <- replace_na(df1$sex, majority_sex)
df1$native.country <- replace_na(df1$native.country, majority_native.country)
colSums(is.na(df1))
```
