
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

-------

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------

1. Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)

```{r}
#Importing libraries and data. Source: https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)

df <- read_csv('Billionaires Statistics Dataset.csv')
df
```

```{r}
#When comparing the richest man with the richest woman, they are 74 and 69 years old. The richest people in the world are mostly seniors between 60 to 80. Furthermore, this graph indicates that there are far more male billionaires. The top 10 richest people in the world are all male, and there are more points in the graph on the right. Note (the finalWorth is in millions)

df %>% ggplot()+ 
    geom_point(mapping=aes(x=age, y= finalWorth))+
  facet_wrap(~gender)
```

2. Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)

```{r}
#Converting the class of selfMade from logical to character.
df$selfMade <- df$selfMade %>% ifelse('True','False')
df %>% head %>% select(selfMade)

```


```{r}
#Using ggplot's facet_wrap function, the majority of male billionaires appear to be self made while the majority of female billionaires are not self made. The world's richest man and the world's richest woman are not self made. For some more detail on the variable `selfMade`, it is defined as someone who achieved their wealth without a large inheritance of wealth or status.  
df %>% ggplot()+ 
  geom_point(mapping=aes(x=age, y=finalWorth ))+
  facet_wrap(~gender+selfMade)
```

3. Make a plot using `position ='dodge'`.

```{r}
#The argument `position = 'dodge'' is best used for comparing graphs of data side by side rather than overlapping. In comparison to fill, dodge makes it easier to compare quantities of variable categories.The most applicable plot to make for either fill or dodge is a bar plot. When comparing the industries of male and female billionaires in the U.S., the industries with the most male billionaires are finance & investments, and technology. The industries with the most female billionaires are food & beverage, fashion & retail, and technology. 

df %>% filter(country == 'United States') %>% ggplot()+ 
  geom_bar(mapping=aes(x=gender, fill=industries), 
           position = 'dodge')

```


4. Make a plot using `position ='fill'`. 

```{r}
#The barplot for self made billionaires by gender gives a much better illustration than the second graph of the proportion of self made billionaires between males and females. 

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=gender, fill=selfMade), 
           position = 'fill')

```


5. Make a plot using `geom_col`.

```{r}
#Interestingly, there is a very high number of billionaires born in the month of January in comparison to the other months. This behavior is strange since the number of billionaires born in January doubles even the second highest month. 

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=birthMonth))

```


6. Change the theme of one of your plots. 

```{r}
#There aren't many themes to chose from, but the linedraw theme makes the graph easier to read.
library(ggthemes)

df %>% ggplot()+ 
  geom_bar(mapping=aes(x=birthMonth))+
  theme_economist()
```


7. Make a plot using combo filter + ggplot

```{r}
df %>% filter(industries == 'Technology'|
                industries == 'Finance & Investments '|
                industries == 'Manufacturing') %>% 
  ggplot()+
  geom_point((aes(x=age, y=finalWorth)), color = 'darkgreen')+
  labs(title = "Billionaires in 3 Industries",
       caption = "Billionaires are compared from the Technology, 
       Finance & Investments, and Manufacturing are compared",
      tag = "Figure 1",x = "Age",y = "Final Worth",
      color = "Gears"
    )+theme_igray()
```


8. Make a plot using combo group_by + summarise + geom_col

```{r}
df %>% group_by(industries, selfMade) %>% 
  summarise(mean_worth = mean(finalWorth, na.rm = TRUE)) %>% 
  ggplot()+
  geom_col(aes(x=selfMade, y=mean_worth, fill=industries), position = 'dodge')
  
```


9. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
[Link to the plot](abc.png)
```

10. In this question, we will use  the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv)

- [Alternative Link](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv')
```



`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))
```

Do the follows. 

  - Use `geom_point` to compare the vaccines used by date in Argentina in 2022. 
  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Notice that the above codes could be piped in one 

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```


Do the follows. 

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021

