
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 6: Visualization - Aesthetic Mapping"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment6.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

### Questions on Your Selected Data

Choose a dataset and produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. 

-------

1. For one continuous variable:

- Density

- Histogram 

- Boxplot

```{r}
#https://www.kaggle.com/datasets/mohidabdulrehman/ultimate-car-price-prediction-dataset
library(tidyverse)
library(dplyr)
library(ggplot2)
df <- read_csv('cars.csv')
df
```
```{r}
#The density plot for the variable 'Mileage' demonstrates that there is a high concentration of cars with 60,000-100,000 miles in the secondary market. There are practically 0 cars in the market with a mileage of over 300,000 miles.

df %>% ggplot()+
  geom_density(mapping = aes(x = Mileage))
```

```{r}
#The histogram of the variable `Mileage` reinforces the pattern observed from the density plot. There is a higher frequency of cars on the secondary market with a mileage between 60,000 and 100,000 miles. There are no cars on the market after 500,000 miles.

df %>% ggplot()+
  geom_histogram(mapping = aes(x = Mileage))
```
```{r}
#The boxplot for the variable `Mileage` indicates the average mileage for a car on the used car market is approximately 80,000 miles (from the vertical line). From the width of the white box, the interquartile range is approximately 60,000 to 120,000 miles. Most of the outliers exceed 120,000 miles.

df %>% ggplot()+
  geom_boxplot(mapping = aes(x = Mileage))
```
2. For one categorical variable

- Barplot

```{r}
#The barplot indicates that there are roughly 3,000 more manual vehicles in the dataset compared to automatic.

df %>% ggplot()+
  geom_bar(mapping=aes(x=`Transmission Type`))
```

-------

3. For two continuous variables

- Scatter Plot

- Line Plot

- Smooth Plot


```{r}
#The upward trend in the scatter plot for the variables `Mileage` and `Price` indicates that these variables are correlated. More specifically, there is a negative correlation between `Mileage` and `Price`. As the miles increase, the points become less dense. However, there are some outliers with 999,999 miles.
max(df$Mileage)
df %>% ggplot()+geom_point(aes(x= Mileage, y=Price))
```

```{r}
df %>% ggplot()+geom_line(aes(x=`Mileage`, y=`Price`))

```
```{r}

df %>% ggplot()+geom_smooth(aes(x=`Mileage`, y=`Price`))
```

-------

4. For one continuous + one categorical variables

- Density

- BoxPlot

```{r}
df %>% ggplot()+
  geom_density(mapping = aes(x = `Mileage`, color = `Transmission Type`))
```
```{r}
df %>% ggplot()+
  geom_boxplot(mapping = aes(x = Mileage, y = `Transmission Type`))
```


-------

5. For two categorical variables: barplot

```{r}
df %>% ggplot()+
  geom_bar(mapping=aes(x= `Transmission Type`, fill= `Body Type`), position = 'fill')
```


-------

6. A plot that contains three variables

```{r}
ggplot(df, aes(x = `Model Year`, y = `Engine Capacity`, size = Price)) +
       geom_point(alpha = 0.7) +
       scale_size_area()
```

