
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 

1. List all the names of the columns of the data

```{r}
library(tidyverse)
df <- read_csv('imdb_top_1000.csv')
names(df)
```
2. Which movies have the highest money earned (Gross)?
```{r}
df %>%
  group_by(Series_Title) %>% 
  summarise(Gross) %>% 
  arrange(-Gross) %>% 
  head()
  
```
3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.
```{r}
df %>%
  group_by(Series_Title) %>% 
  summarise(IMDB_Rating) %>% 
  arrange(IMDB_Rating) %>% 
  head()
```
4. Which year have the most number of movies released in the list?  What is the total of money earned on that year? Ans: 2014 ; total revenue = $2,755,629,221	
```{r}
df %>% 
  group_by(Released_Year) %>% 
  count(Released_Year) %>% 
  arrange(-n)
```
```{r}
df %>%
  filter(Released_Year == '2014') %>% 
  summarise(total_revenue = sum(Gross,na.rm=TRUE))
```

5. What is the total money earned per movies on average? Ans: $68,034,751
```{r}
df %>% 
  summarise(avg_revenue = mean(Gross,na.rm=TRUE))
```
6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9. Ans: 1,981,739	
```{r}
df %>% 
  group_by(Released_Year) %>% 
  summarise(avg_votes = mean(No_of_Votes,na.rm=TRUE))
```
```{r}
df %>%
  filter(IMDB_Rating>9) %>% 
  summarise(avg_votes = mean(No_of_Votes,na.rm=TRUE))

```
7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile.
```{r}
df2020 <- df %>% filter(Released_Year == '2020')
summary(df2020$No_of_Votes)
```
```{r}
#There is only one movie from 2020 that falls within the range of votes.
#Hence, the average Meta score only represents one movie.

df2020 %>% 
  filter(No_of_Votes >= 84127 & No_of_Votes < 106083) %>% 
  summarise(avg_meta_score = mean(Meta_score,na.rm=TRUE))

```
```{r}
df2020
```

8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  
```{r}
df$Runtime <- df$Runtime %>% 
  str_remove('min')

df <- df %>%
  mutate(Runtime = as.numeric(Runtime,na.rm = TRUE))
head(df)
```
```{r}
df %>% 
  filter(Released_Year >=2010 & Released_Year <=2019) %>% 
  summarise(avg_run_time_2010s = mean(Runtime, na.rm = TRUE))
```
```{r}
cor(df$Runtime, df$IMDB_Rating, use = "complete.obs")
```

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=FALSE}
# Select only character columns
df %>% select_if(is.character)

# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```
10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange

```{r}
#Removing non-useful columns from the dataframe.
df <- df %>% select(-Poster_Link, -Certificate, -Overview)
names(df)
```
```{r}
#Filtering movies by the big category "Action, Crime, Drama"
df %>% filter(Genre == "Action, Crime, Drama")
```
```{r}
d
```

