
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 5 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 5 - Extra Credits. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

1. Download the c2015 dataset to your computer at [this link](../data/c2015.xlsx). Load the library `readxl` (library(readxl)) then use the function `read_excel()` to read the c2015 dataset.  The data is from Fatality Analysis Reporting System (FARS). The data includes vital accidents information, such as when, where, and how the accident happened. FARS also includes the drivers and passengers' information, such as age,gender etc.  Some of the fatal accident had multiple vehicles involved.  More information about FARS can be found at: https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars

```{r}
#Importing Libraries and importing Excel data with read_excel

library(tidyverse)
library(dplyr)
library(readxl)
df <- read_excel('c2015.xlsx')
head(df)
```


2. Let's study the variable `SEX`. How many missing values in the NAs form? Ans: 0  in NA's form, however, there are two categories, Not Rep and Unknown.
```{r}
#Checking the number of missing values in NA form within the variable 'SEX'.

sum(is.na(df$SEX))
```
```{r}
#Checking the categories within the variable 'SEX'. There are two categories for missing values.

table(df$SEX)
```

3. Still with variable `SEX`. There are missing values in this variables that are not `NAs`. Identify the forms of missing values in this variable. Change all the forms of missing values to `NAs`. Ans: As mentioned in question two there are two forms of missing values, 'Not Rep' and 'Unknown'.
```{r}
#Using 'na_if' to replace values with NA

df$SEX <- df$SEX %>% na_if('Unknown') %>% 
  na_if('Not Rep')
table(df$SEX)

```

4. Still with variable `SEX`. After all the missing values are in the `NAs` form.  Change the missing values of this variable to the majority sex.
```{r}
#Since the majority of participants were male, the number of male participants increased after this change while the number of female participants remained the same.

majority_sex = names(which.max(table(df$SEX)))
df$SEX <- df$SEX %>% replace_na(majority_sex)
table(df$SEX)
```
```{r}
#Checking for missing values after the changes are made to the variable.

sum(is.na(df$SEX))
```

5. Let's study variable `AGE`. Use the `table` function to check out the values of these variable and forms of missing values. Use `na_if` to change all the forms of missing values to `NAs`.
```{r}
#Using the 'table' function to explore the variable 'AGE' and find missing values.

table(df$AGE)
```
```{r}
#Replacing the category 'Unknown' with NA inside AGE.

df$AGE <- df$AGE %>% na_if('Unknown')
sum(is.na(df$AGE))
```
```{r}
#Using table to observe the changes made to the variable 'AGE'.

table(df$AGE)
```

6. Still with variable `AGE`.  Use the `str_replace` to replace `Less than 1` to '0' (character 0, not number 0).
```{r}
#Using the 'str_replace' function to replace the category 'Less than 1 within 'AGE' to the string 0.

df$AGE <- df$AGE %>% str_replace('Less than 1', '0')
```

7. Still with variable `AGE`. Use the `class` function to check the type of this variable.  Use the `as.numeric` function to change the type of the variable to numeric.

```{r}
#Viewing the class of the variable 'AGE'. 
class(df$AGE)
```
```{r}
#Converting the class of AGE from character to numeric; the changes are saved to the dataframe.

df <- df %>% mutate(AGE = as.numeric(AGE,na.rm = TRUE))
class(df$AGE)
table(df$AGE)
```

8. Still with variable `AGE`. Replace the missing values `NAs` by the mean of the variable. 
```{r}
#Creating a variable for the average age of the participant, and then using the 'replace_na' function to change the missing values to the average age. The changes are saved to the dataframe.

mean_age = mean(df$AGE, na.rm = TRUE)
df$AGE <- df$AGE %>% replace_na(mean_age)
sum(is.na(df$AGE))
```
```{r}
#Observing the values within the variable 'AGE' after the changes were made.

table(df$AGE)
```

9. Let's fix the variable `TRAV_SP`. Do the follows.

- Use the `table` function to check all the values of this variable.  Use the `str_remove` to remove the `MPH` in each value.
- Remove the row with the travel speed `Greater`
- Use the `str_replace` function to replace `Stopped` by '0' (dont forget the quotation mark around 0)
- Use `na_if` to change all the forms of missing values to `NAs`
- Check the type of this variable using `class`. Use `as.numeric` to change the type to numeric.
- Use `replace_na` to replace the `NAs` to the median of the variable. 

```{r}
#Checking all the values within the variable 'TRAV_SP' with the 'table' function.

table(df$TRAV_SP)
```
```{r}
#Removing the string 'MPH' next to the values in the variable TRAV_SP so the class can be changed in a later step.

df$TRAV_SP <- df$TRAV_SP %>% str_remove('MPH')
sum(table(df$TRAV_SP))
```
```{r}
#Removing the row that produced a value of 'Greater' in the column 'TRAV_SP'. The subset function was used, but I'm sure there are many other methods that can be applied.

df <- df %>% subset(TRAV_SP != 'Greater')
```
```{r}
#Using table to observe the changes made to the dataframe.

table(df$TRAV_SP)
```
```{r}
#Replacing the string 'Stopped' with the character 0 in the variable 'TRAV_SP'. The changes are then observed with the table function.

df$TRAV_SP <- df$TRAV_SP %>% 
  str_replace('Stopped','0')
table(df$TRAV_SP)
```
```{r}
#Replacing the values 'Not Rep' and 'Unknown' to NA within the variable 'TRAV_SP'.

df$TRAV_SP <- df$TRAV_SP %>% na_if('Not Rep') %>% 
  na_if('Unknown')
table(df$TRAV_SP)
```
```{r}
#Viewing the class of the variable 'TRAV_SP'.

class(df$TRAV_SP)
```
```{r}
#Converting the class of the variable 'TRAV_SP' from character to numeric. The changes are saved to the dataframe.

df <- df %>% mutate(TRAV_SP = as.numeric(TRAV_SP,na.rm = TRUE))
class(df$TRAV_SP)
```
```{r}
#Assigning a variable for the median speed of travel. All the NAs in the variable 'TRAV_SP' are replaced by this value.

median_trav_sp = median(df$TRAV_SP, na.rm = TRUE)
df$TRAV_SP <- df$TRAV_SP %>% replace_na(median_trav_sp)
sum(is.na(df$TRAV_SP))
```
```{r}
#Viewing the first 6 rows of the dataframe.
head(df)
```
10. Find the correlation between Age of the drivers and Travel speed (`TRAV_SP`). Hint:  You want to look at the seat positions (`SEAT_POS` variable) to filter out the observations about the drivers, then calculate the correlation. ANS: The correlation is approximately -0.1186 meaning that younger drivers are positively correlated with higher speeds of travel.

```{r}
#Creating a subset dataframe for only drivers with the function 'filter'. The correlation between the variables 'AGE' and 'TRAV_SP' is taken from the resulting dataframe.

driver_df <- df %>% filter(SEAT_POS == 'Front Seat, Left Side')
cor(driver_df$AGE,driver_df$TRAV_SP)
```